### 2010-07-02 paper/IIWAS2010/content/introduction.tex
latest working copy
latest version

When traveling though a jungle to a destination, it is easy to get lost.  The first person to journey somewhere may make a number of mistakes when trying to find the best path to their destination. Those who come later find it easier to reach the destination if a well-marked trail has been created. Olsen and Malizia describe this idea as \emph{exploiting trails} \cite{5379671}.  Rather than treating a user's discovery experience as a unique entity, one can exploit the fact that a similar search may have already been performed.  In one study, almost 40 percent of all queries were repetitions of previous queries\cite{1277770}. Thus, reuse of prior searches is one way to optimize the search
process.  Morpheus is a question answering system motivated by reuse of prior web search pathways to yield an answer to a user query. Morpheus follows path-finders to their destinations and not only marks the trail, but also provides a taxi service to take followers to similar
destinations.

@@

Morpheus represents user-query details in a semi-structured query (SSQ). It assumes the terms belong to classes of a \textit{consistent} realm-based ontology, that is, one having a singly rooted heterarchy whose subclass/superclass relations have meaningful semantic interpretations. When a path-follower enters a query, Morpheus ranks SSQs in the store based on class similarity. Suppose a regular user asks -\textit{ a 1997 Toyota Camry V6 needs what size tires?} In this query the classes associated with terms, e.g. \emph{Manufacturer} with \emph{Toyota}, helps us identify similar queries.

>>>>

When traveling though a jungle to a destination, it is easy to get lost.  The first person to journey somewhere may make a number of mistakes when trying to find the best path to their destination. Those who come later find it easier to reach the destination if a well-marked trail has been created. Olsen and Malizia describe this idea as \emph{exploiting trails} \cite{5379671}.  Rather than treating a user's discovery experience as a unique entity, one can exploit the fact that a similar search may have already been performed.  In one study, almost 40 percent of all queries were repetitions of previous queries \cite{1277770}. Thus, reuse of prior searches is one way to optimize the search
process.  Morpheus is a question answering system motivated by reuse of prior web search pathways to yield an answer to a user query. Morpheus follows path-finders to their destinations and not only marks the trail, but also provides a taxi service to take followers to similar destinations.

Morpheus focuses on the deep (or hidden) web to answer questions because its scale and database driven property implies the existence of huge stores of quality information \cite{1453931}.  Web forms act as an interface to this information.  Morpheus employs user exploration through these web forms to learn the types of data each deep web location provides.

@@

Morpheus represents user-query details in a semi-structured query (SSQ). It assumes the query terms belong to classes of a \textit{consistent} realm-based ontology, that is, one having a singly rooted heterarchy whose subclass/superclass relations have meaningful semantic interpretations. When a path-follower enters a query, Morpheus ranks SSQs in the store based on class similarity. Suppose a path follower asks -\textit{ a 1997 Toyota Camry V6 needs what size tires?} In this query the classes associated with terms, e.g. \emph{Manufacturer} with \emph{Toyota}, help us identify similar queries.

### 2010-07-02 paper/IIWAS2010/content/related_work.tex
latest version

Several community QA systems have been developed, such as Yahoo! Answers \cite{yahooanswers2008}. These sites enlist a community of users to answer questions posed by other users.  This idea has led to socially integrated QA systems such as Aardvark \cite{vark2010}. In addition to having questions answered by users, Aardvark can contact potential question answerers through media beyond web pages (e.g. instant messages, mobile phones).  This allows for faster responses to questions.  Morpheus relies on both its user interface and previous similar questions answered by a community of users to give a fast reliable answer.

@@

The DBpedia\footnote{http://dbpedia.org} project is based on a community of contributors extracting semantic information from the Wikipedia and making it available on the Web.  Wikipedia semantics includes info-box templates, categorization information,  images, Geo-coordinates, links to external Web pages, disambiguation pages, and redirects between pages in Wiki markup form \cite{Bizer2009}.  DBpedia does not define any new relations between the Wikipedia categories.

@@

YAGO is a semi-automatically constructed ontology from the Wikipedia pages, info-boxes, categories and WordNet\footnote{http://wordnet.princeton.edu} synsets heterarchy \cite{Suchanek2009phd}. YAGO uses the Wikipedia page titles as its ontology individuals and categories as its ontology classes. YAGO uses only the nouns from WordNet and ignores the WordNet verbs and adjectives.  YAGO discovers connections between a WordNet synsets and a Wikipedia categories parsing the category names and matching the parsed category components with the WordNet synsets. Those Wikipedia categories having no WordNet match are ignored in the YAGO ontology.  The ontology's heterarchy is built using the \textit{hypernym} and \textit{hyponym} relations of the WordNet synsets.

We have attempted to use similar techniques to construct ontologies for question answering realms.  Thus far, these ontologies can be use to classifiy terms, but their classes do not appropriately categorize deep web query parameters. Therefore we have been relegated to manually generating ontologies.  % TODO - add "this is under investigation"

>>>>

Several community QA systems have been developed, such as Yahoo! Answers \cite{yahooanswers2008}. These sites enlist a community of users to answer questions posed by other users.  This idea has led to socially integrated QA systems such as Aardvark \cite{vark2010}. In addition to having questions answered by users, Aardvark can contact potential question answerers through media beyond web pages (e.g. instant messages, mobile phones).  This allows for faster responses to questions.  Morpheus relies on both its user interface and referencing similar questions previously answered.

@@

The DBpedia\footnote{http://dbpedia.org} project is based on a community of contributors extracting semantic information from the Wikipedia and making this information available on the Web.  Wikipedia semantics includes info-box templates, categorization information,  images, geo-coordinates, links to external web pages, disambiguation pages, and redirects to pages in Wiki markup form \cite{Bizer2009}.  DBpedia does not define any new relations between the Wikipedia categories.

@@

YAGO is a semi-automatically constructed ontology obtained from the Wikipedia pages, info-boxes, categories, and WordNet\footnote{http://wordnet.princeton.edu} synsets heterarchy \cite{Suchanek2009phd}. YAGO uses the Wikipedia page \emph{titles} as its ontology \emph{individuals} and \emph{categories} as its ontology \emph{classes}. YAGO uses only the nouns from WordNet and ignores the WordNet verbs and adjectives.  YAGO discovers connections between WordNet synsets and Wikipedia categories, parsing the category names and matching the parsed category components with the WordNet synsets.  Wikipedia categories having no WordNet match are ignored in the YAGO ontology.  The ontology's heterarchy is built using the \textit{hypernym} and \textit{hyponym} relations of the WordNet synsets.

We use YAGO's principles to construct ontologies that provide similarity measures for answering questions within the same domain.  Thus far, these ontologies can be used to classifiy terms, however their classes do not always appropriately categorize query parameters.  Therefore, we are researching XXX in order to XXX.% TODO - fill in the XXX

### 2010-07-02 paper/IIWAS2010/content/system_arch.tex
latest version

To build a corpus for each of the leaf nodes in the ontology, we extract \emph{terms} from the Wikipedia pages associated with the DBpedia categories found in its blanket. From this term corpus, we can find the likelihood of a term belonging to a particular category. This assists in categorizing terms in a \emph{path-follower} query. The likelihood is determined by the probability of a category given a term using \textit{Bayes Rule} (Eq. \ref{eq:bayesrule}), since we can easily obtain the term-category and term-corpus probabilities as relative frequencies.

@@

\begin{equation}
\label{eq:bayesrule}
P (category | term) = \frac{P(term | category) P(category)}{P(term)}
\end{equation}

@@

The \emph{Query Resolution Recorder} (QRR) is an instrumented web browser that records the interactions of a path-finder answering a question. The path-finder also uses the tool to identify ontological categories associated with search terms. Morpheus stores the query together with its terms and categories as an SSQ.  Table \ref{tab:ssq_example} is an example showing the SSQ model of the query ``A 1997 Toyota Camry V6 has what size tires?". The SSQ in Table \ref{tab:ssq_example} is said to be \emph{qualified} because the classes associated with its terms have been identified. Using the QRR, the path-finder is also able to identify where answers can be found within traversed pages.

@@

Queries inserted by path-followers take a different approach. The Morpheus search process parses and tags queries in order to record important terms. The system assigns the most probable realm given the terms in the query as calculated from realm-specific copora. Once the realm is assigned, an ontology search is performed to assign classes to the terms. An SSQ is constructed and the system attempts to match this candidate SSQ to existing QRM SSQs. Rather than matching exact query terms, the system matches input and output classes, because a QRM can potentially answer similar queries.

@@

The \emph{class divergence} of two classes within the ontology characterizes their similarity.  This solution is motivated by the concept of multiple dispatch in CLOS and Dylan programming for generic function type matches.

We consider the category match as a type match and we use \emph{category divergence} to calculate the relevance between the candidate SSQ and a qualified SSQ. Each qualified SSQ will have input terms, output terms, and associated classes, and one realm (from the QRM).  For the candidate SSQ, the relevant categories for terms are determined from the NLP engine and corpora. The calculation of a realm for a candidate query is performed using the n-grams found within the query and any probabilities as found with $p(realm|n\-gram)$. %% TODO - specify freq and reference joirdans stuff in section 3.2
We match QRMs that belong to the same realm of the candidate SSQ. The relevance of a qualified SSQ to the candidate SSQ is determined by aggregating the divergence measure of input term classes associated with each SSQ. In addition, we order QRMs in the data store 
by decreasing relevance. The order provides a ranking for the results 
to the user. The following describes \emph{category divergence} 
in detail.

@@

We employ a quasimetric, \textit{category divergence (cd)}, between a source category and a target category using the \textit{topological structure} of the categories in an ontology. We write $S \prec T$ for the reflexive transitive closure of the supercategory relation. Let $d(P,Q)$ represent the hop distance in the directed ontology inheritance graph from $P$ to $Q$. The divergence between a source and target category ranges from zero (for identical categories) to one (for type incompatible categories). Let $S$ be the source category, $T$ be the target category, and $C$ be a least common ancestor category of $S$ and $T$ i.e., one that minimizes $d(S,C) + d(T,C)$. The category divergence between $S$ and $T$ is defined given by:

@@

Note, if $S \prec T$ and $S \not\prec Q$ then $cd(S,T) < cd(S,Q)$, that is, the divergence of a source category to a target ancestor category is smaller than the divergence of a source category to any category that is not an ancestor. This is an important property in determining the compatibility of categories for answering queries.  If a SSQ answers queries concerning an ancestor category, it is more relevant that a SSQ that answers queries from any non-ancestral category.

Suppose we want to find the category divergence between \textit{Bus} and \textit{Sedan} (Figure \ref{fig:vehicular_ontology}). \textit{Sedan} is a subcategory of \textit{Automobile}, and \textit{Bus} is a subcategory of \textit{Land\_Vechicle}. Therefore, the least common ancestor ($C$) for these two is \textit{Land\_Vechicle}. In addition, the tree height \textit{h = 4}. So, the normalized divergence $cd$\textit{(Bus, Sedan)} is \textit{0.5} that is calculated from \textit{d(Bus, Root) = 3}, \textit{d(Bus, Land\_Vehicle) = 1}, and \textit{d (Sedan, Land\_Vehicle) = 2}.

@@

Once we have relevant QRMs through QRM ranking for a given user query, we can produce answers by re-running the pathways.

@@

%When a user submits a new query to the Morpheus system, one or more stored QRMs are executed to obtain results. A QRM contains an XML script representing an algorithm that will obtain the required result. The evaluation of this script by the QRE simulates the behavior of a human browsing the web, that is, clicking buttons to follow links, submitting forms, highlighting data, cutting and pasting text, and constructing an answer in the form of a text string. There is greater complexity in dealing with web forms than with links. At present, the GET form submission method is used. This means that in preparation to submit a form, the Morpheus System builds a querystring consisting of name-value pairs. For a given form, there may be text fields, drop-down boxes, and among other components. In order to support form submission in such circumstances, the QRE performs different actions based upon the input type. For text fields, simply adding a key/value pair to the query string suffices. For a drop-down box, the QRE extracts the form element to find the appropriate value to include in the query string. It does this by examine each option in the drop-down and matching the user’s input to the text field in the option. The value attribute of the option is then added to the querystring. The value and text attributes of a dropdown option might be the same, but they are not guaranteed to be. All input data are kept in an internal table. The QRM script contains keys that allow the QRE to map SSQ inputs from the table to the appropriate form inputs. Furthermore, highlighted text segments collected during execution of the script are stored in the table. Once all actions in the script have been carried out, the QRE returns the textual answer this process has created. This answer is then be rendered in a browser displayed to the user.

>>>>

To build a corpus for each of the leaf nodes in the ontology, we extract \emph{terms} from the Wikipedia pages associated with the DBpedia categories found in its blanket. From this term corpus, we can find the likelihood of a term belonging to a particular class. This assists in classifying terms in a \emph{path-follower} query. The likelihood is determined by the probability of a class given a term using \textit{Bayes Rule} (Eq. \ref{eq:bayesrule}), since we can easily obtain the term-class and term-corpus probabilities as relative frequencies.

@@

\begin{equation}
\label{eq:bayesrule}
P (class | term) = \frac{P(term | class) P(class)}{P(term)}
\end{equation}

@@

The \emph{Query Resolution Recorder} (QRR) is an instrumented web browser that records the interactions of a path-finder answering a question. The path-finder also uses the tool to identify ontological classes associated with search terms. Morpheus stores the query together with its terms and classes as an SSQ.  Table \ref{tab:ssq_example} is an example showing the SSQ model of the query ``A 1997 Toyota Camry V6 has what size tires?". The SSQ in Table \ref{tab:ssq_example} is said to be \emph{qualified} because the classes associated with its terms have been identified. Using the QRR, the path-finder is also able to identify where answers can be found within traversed pages.

@@

Queries inserted by path-followers take a different approach. The Morpheus search process parses and tags queries in order to record important terms. The system assigns the most probable realm given the terms in the query as calculated from realm-specific copora. Once the realm is assigned, an ontology search is performed to assign classes to the terms. An SSQ is constructed and the system attempts to match this candidate SSQ to existing QRM SSQs. Rather than matching exact query terms, the system matches input and output classes, because a QRM can potentially answer many similar queries.

@@

The \emph{class divergence} of two classes within the ontology characterizes their dissimilarity.  This solution is motivated by the concept of multiple dispatch in CLOS and Dylan programming for generic function type matches.

We consider the class match as a type match and we use \emph{class divergence} to calculate the relevance between the candidate SSQ and a qualified SSQ. Each qualified SSQ will have input terms, output terms, and associated classes, and one realm (from the QRM).  For the candidate SSQ, the relevant classes for terms are determined from the NLP engine and corpora. The calculation of a realm for a candidate query is performed using the n-grams found within the query and any probabilities as found with $p(realm|n\-gram)$. %% TODO - specify freq and reference joirdans stuff in section 3.2
We match QRMs that belong to the same realm of the candidate SSQ. The relevance of a qualified SSQ to the candidate SSQ is determined by aggregating the divergence measure of input term classes associated with each SSQ. In addition, we order QRMs in the data store 
by decreasing relevance. The order provides a ranking for the results 
to the user. The following describes \emph{class divergence} 
in detail.

@@

We employ a quasi-metric, \textit{class divergence (cd)}, between a source class and a target class using the \textit{topological structure} of the classes in an ontology. We write $S \prec T$ for the reflexive transitive closure of the superclass relation. Let $d(P,Q)$ represent the hop distance in the directed ontology inheritance graph from $P$ to $Q$. The divergence between a source and target class ranges from zero (for identical classes) to one (for type incompatible classes). Let $S$ be the source class, $T$ be the target class, and $C$ be a least common ancestor class of $S$ and $T$ i.e., one that minimizes $d(S,C) + d(T,C)$. The class divergence between $S$ and $T$ is defined given by:

@@

Note, if $S \prec T$ and $S \not\prec Q$ then $cd(S,T) < cd(S,Q)$, that is, the divergence of a source class to a target ancestor class is smaller than the divergence of a source class to any class that is not an ancestor. This is an important property in determining the compatibility of classes for answering queries.  If an SSQ answers queries concerning an ancestor class, it is more relevant that an SSQ that answers queries from any non-ancestral class.

Suppose we want to find the class divergence between \emph{Bus} and \emph{Sedan} from the ontology shown in Figure \ref{fig:vehicular_ontology}. \emph{Sedan} is a subclass of \emph{Automobile} which is a subclass of \emph{Land\_Vehicle}.  \emph{Bus} is a subclass of \emph{Land\_Vehicle}, which is their least common ancestor $C$.

The longest path to the tree root is $h = 4$. Therefore, the normalized divergence $cd(Bus, Sedan)$ is $0.5$. The value is calculated from $d(Bus, Root) = 3$, $d(Bus, Land\_Vehicle) = 1$, and $d(Sedan, Land\_Vehicle) = 2$.

@@

Once we have ranked the QRMs for a given user query, we can produce answers by re-running the pathways stored in the QRMs.

@@
